Statistical Inference
========================================================
author: Jordan DeHerrera
date: 11/4/2015

### PDF, PMF, CDF, and Survival Function

#### Probability Mass Function

A probability mass function is the probability of a specific value occurring.

* Mutually exclusive values (e.g., brunette, blonde, or redhead)
* Can be used for experiment results (e.g., success or failure)

#### Coin Flip Example of a Binomial Probability Function

* Let $X$ be the result of a coin flip where $X=0$ represents
tails and $X = 1$ represents heads.

$$
p(x) = (1/2)^{x} (1/2)^{1-x} ~~\mbox{ for }~~x = 0,1
$$

* Suppose that we do not know whether or not the coin is fair; Let
$\theta$ be the probability of a head expressed as a proportion
(between 0 and 1).

$$
p(x) = \theta^{x} (1 - \theta)^{1-x} ~~\mbox{ for }~~x = 0,1
$$

#### Probability Density Function

A probability density function is the probability for a range of values occurring.

* Any value on a continuum (e.g., height and weight)
* The area under the line in a graph
* Use area calculations to determine the probability density of two values

#### Call Center Example of PDF

The following example shows how a probability density function works for a very simple (and easily calculated density).
* We are given the following function for a call center's performance:
$$
f(x) = \left\{\begin{array}{ll}
    2 x & \mbox{ for } 1 > x > 0 \\
    0                 & \mbox{ otherwise} 
\end{array} \right. 
$$

#### Visual Demonstration of PDF

* This is how we can visualize the shape of the distribution (note how it forms a perfect right triangle).
* The area of the right triangle can be calculated as $$1/2 * b * h$$

```{r, fig.height = 5, fig.width = 5, echo = TRUE, fig.align='center'}
x <- c(-0.5, 0, 1, 1, 1.5); y <- c( 0, 0, 2, 0, 0)
plot(x, y, lwd = 3, frame = FALSE, type = "l")
```

#### Problem Statement

Suppose the operator of the call center wanted to know the probability that their call center answered 75% or fewer of the calls?  This can be represented by the following graph:

```{r, fig.height = 5, fig.width = 5, echo = FALSE, fig.align='center'}
plot(x, y, lwd = 3, frame = FALSE, type = "l")
polygon(c(0, .75, .75, 0), c(0, 0, 1.5, 0), lwd = 3, col = "lightblue")
```

#### Solution

To solve the problem, we just need to calculate the area of the triangle (the probability density function since we're looking for the probability of a distribution rather than a discrete value).

```{r}
1.5 * .75 / 2
pbeta(.75, 2, 1)
```

#### Cumulative Distribution Function and Survival Function


##### Probability Distribution Function

The probability of a distribution being less than a certain value.

##### Survival Function

The probability of a distribution being greater than a certain value.
